---
title: "Lab4"
author: "Karthikcharan Suresh"
date: "10/24/2019"
output: html_document
---
Github username: ksuresh1001

Repository link : https://github.com/ksuresh1001/ds202_lab4

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading Data: Cyclone Football 

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```

```{r}
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
```


## Part one: Cleaning data
# 1.
```{r}
library(dplyr)
library(tidyr)
offense$Name<-as.factor(offense$Name)
offense$Opponent_Opponent<-as.factor(offense$Opponent_Opponent)
defense$Opponent_Opponent<-as.factor(defense$Opponent_Opponent)
defense$Name<-as.factor(defense$Name)
```

# 2.
```{r}
bio$Weight<-as.numeric(bio$Weight)
colo<-c("Receiving_REC",
        "Receiving_YDS",
        "Receiving_TD",
        "Rushing_ATT",
        "Rushing_YDS",
        "Rushing_TD",
        "Passing_CMP-ATT",
        "Passing_YDS",
        "Passing_TD",
        "Passing_INT")
offense[colo]<-sapply(offense[colo], as.numeric)
cold<-c("Tackles_Solo","Tackles_ASST","Tackles_TFL","Tackles_Sack","Turnover_FF","Turnover_FR","Turnover_INT","Pass_QBH","Pass_PB")
defense[cold]<-sapply(defense[cold],as.numeric)
```

# 3.
```{r}
bio<-separate(data=bio,col=Height,
              into=c("Feet","inc"),
              sep="\\-")
bio$Feet<-as.numeric(bio$Feet)
bio$inc<-as.numeric(bio$inc)
bio$height_in_inches<-(bio$Feet*12+bio$inc)
head(bio)
colb<-c("Name",
        "Position",
        "Class",
        "Hometown",
        "Highschool")
bio[colb]<-lapply(bio[colb],factor)
bioClean<-select(bio,-c("Feet","inc"))
head(bioClean)
defClean<-defense
offClean<-offense
```

# 4.
```{r}
sapply(offClean,class)
sapply(defClean,class)
sapply(bioClean,class)
```

## Part 2 Tidying Part A
# 2.1.
```{r}
defCleanGather = gather(defClean,key="Statistic", value="Value", 3:11)
```

# 2.2.
```{r, echo=FALSE}
library(ggplot2)
ggplot(defCleanGather, aes(x=Value)) + geom_histogram(binwidth=0.8) + facet_wrap(~Statistic)
```

The rarest defensive skills is the turnover skills because most of those values are 0 and only a few are 1

# 2.3.
```{r, echo=FALSE}
WVKSsolo <- defCleanGather %>%
  filter(Statistic=="Tackles_Solo") %>% 
  spread(key = Opponent_Opponent,
         value = Value)
ggplot(WVKSsolo, aes(x=`West Virginia`, y=`Kansas State`)) + geom_point() + ggtitle("Solo Tackles Against West Virginia and Kansas State")
```
```

#We had a much better defense against Kansas State compared to west virginia as there were large number of solo tackles against Kansas State


## Part 2 Tidying Part B
# 1.
```{r}
bioCleanSeparate <- separate(bioClean, 
                             Hometown, 
                             into = c("City","State"), 
                             sep = ",")
```

# 2.
```{r, echo=FALSE}
StateCount <- bioCleanSeparate %>% 
              group_by(State) %>% 
              summarize(n=n())
```

#How many players are there from each state?: 
Arizona=3, Arkansas=2, Australia=2, California=8, D.C.=1, Florida=8, Georgia=2, Illinois=5, Iowa=44, Kansas=3, Massachusettes=1, Maryland=1, Michigan=5, Minnesota=2, Missippi=1, Montana=1, Nebraska=1, Nigeria=1, Ohio=10, Oklahoma=4, Pennsylvania=2, South Carolina=2, Texas=9, Virginia=1, Wisconsin=4.

## Part three: Joining data frames

# 3.1.

```{r}
offbio<-inner_join(offClean,
                   bioClean,
                   by="Name")
colnames(offbio)
```

# 3.2.

```{r, echo=FALSE}
ggplot(data=offbio,
       aes(x=Weight,
           y=Receiving_YDS))+
           geom_point()+  
           labs(x="Weight",
           y="Receiving Yards",
           title="Scatterplot of Weight and Receving Yards")
```

yes, there is relationship as Players who weigh 225 pounds have the highest number of recieving yards. 

#3.3 How did weight matter when it comes to Passing_YDS

```{r, echo=FALSE}
ggplot(data=offbio,
       aes(x=Weight,
           y=Passing_YDS))+
           geom_point()+  
           labs(x="Weight",
           y="Solo Tackles",
           title="Scatterplot of Weight and Passing Yards")
           